### 2009-06-08 doc/manuals/janitor/tex_installation/installation.tex
Michael's additions and some pending corrections by Steffen.

In order to use the Janitor, the two perl packages listed in
table~\ref{tab:install_dependencies} are required. To have
the WebService interface for the Janitor, the packages listed in the
table~\ref{tab:install_dependencies_optional} need to be installed before
the build process is.

\begin{table}[!h]
   \begin{center}
        \mycaption{Required perl packages for Janitor.}{Log4perl is used for the internal logging of Janitor, while the 
                   Redland RDF libraray is used for accessing the knowledge base of Runtime Environments.}
        \label{tab:install_dependencies}
	\begin{tabular}{|p{3cm}|p{7cm}|}
	\hline
	   liblog-log4perl-perl & \textit{Log4perl is a port of the log4j logging package}\\
	\hline
	   librdf-perl          & \textit{Perl language bindings for the Redland RDF library}\\
	\hline
	\end{tabular} 
   \end{center}
\end{table}
\begin{table}[!h]
   \begin{center}
        \mycaption{Optional libraries for Janitor.}{The library libperl-dev provides header files which are needed to link the 
                   WebService to the Perl interpreter.}
        \label{tab:install_dependencies_optional}
	\begin{tabular}{|p{3cm}|p{7cm}|}
	\hline
	   libperl-dev & \textit{ Perl library: development files}\\
	\hline
	\end{tabular}
   \end{center}
\end{table}
\forcelinebreak

@@

Furthermore it is recommended to install the ontology editor 
Prot\`eg\`e\footnote{\href{http://protege.stanford.edu}{http://protege.stanford.edu}} in order to be to easily maintain the 
knowledge database of installable packages.

@@

The current version of the Janitor can be configured using the common file \textit{arc.conf}. It is expected in the 
configuration directory \textit{etc}. The Janitor is using the environment variable \textit{NORDUGRID\_CONFIG}
to determine the location of the corresponding file. If that variable is not set, the default location \textit{/etc/arc.conf} will be used. 
The configuration is assigned by the section \lbrack janitor\rbrack. The table~\ref{tab:arc_conf_tags} contains the available tags
for janitor configuration.

@@

The parameter \texttt{enable} defines whether the Janitor shall be used within A-REX or not. Use the value \texttt{"0"} to disable Janitor. The
\texttt{uid} and the \texttt{gid} are defining which effective the user and group id shall be used for Janitor. 
The \texttt{registrationdir} describes the directory in which the subdirectories \texttt{jobs} and \texttt{rtes} will be created.
In these directories the states of the jobs and the runtime environments will be stored. 
The knowledge base of installable packages is specified by the parameter \texttt{catalog}. 
Its value can be any kind of URL pointing to a file written in the Resource Description Framework (RDF) format.
The specification of the RDF file will be explained in detail in section~\ref{sec:catalog}.
The parameter \texttt{downloaddir} assigns the directory in which the installation files will be saved after they have been 
downloaded or copied from the repository which was specified by the catalog. The \texttt{installationdir} finally specifies the
directory into which all packages will be installed. Instead of the other directories, the \texttt{installationdir} should be 
available for all computing elements, i.e. by using a shared volume.
If the configuration file furthermore contains the \texttt{runtimedir} tag within the section \texttt{grid-manager}, the Janitor will also
create a symbolic link in the \texttt{runtimedir} pointing to the configuration script of the installation done by the Janitor.
The tags \texttt{jobexpirytime} and \texttt{rteexpirytime} are used for automated cleanup and is defined in seconds.
The default value for the \texttt{jobexpirytime} is seven days and for the \texttt{rteexpirytime} three days. 
The additional tags \texttt{allow\_base} \texttt{deny\_base} \texttt{allow\_rte} and \texttt{deny\_rte} are used to include
or exclude certain base packages or runtime environments of the catalog. This feature is useful, if the catalog is maintained by
a higher organization.
The path to the log4perl configuration file is defined by the tag \texttt{logconf}.
An examples how to configure arc and log4perl is provided in the Listings~\ref{lst:arc_conf} and ~\ref{lst:log_conf} and.

>>>>

The Janitor requires the two perl packages listed in
table~\ref{tab:install_dependencies}. To have
the WebService interface for the Janitor, the packages listed in the
table~\ref{tab:install_dependencies_optional} need to be installed before
the build process is. The Perl modules are available on CPAN and ship
with all major Linux distributions.

\begin{table}[!h]
   \begin{center}
        \mycaption{Required perl packages for Janitor.}{Log4perl is used for the internal logging of Janitor, while the 
                   Redland RDF library is used for accessing the knowledge base of Runtime Environments.}
        \label{tab:install_dependencies}
	\begin{tabular}{|p{3cm}|p{7cm}|}
	\hline
	   liblog-log4perl-perl & \textit{Log4perl is a port of the log4j logging package}\\
	\hline
	   librdf-perl          & \textit{Perl language bindings for the Redland RDF library}\\
	\hline
	\end{tabular} 
   \end{center}
\end{table}
\begin{table}[!h]
   \begin{center}
        \mycaption{Optional libraries for Janitor.}{The library libperl-dev provides header files which are needed to link the 
                   WebService to the Perl interpreter.}
        \label{tab:install_dependencies_optional}
	\begin{tabular}{|p{3cm}|p{7cm}|}
	\hline
	   libperl-dev & \textit{ Perl library: development files}\\
	\hline
	\end{tabular}
   \end{center}
\end{table}
\forcelinebreak

@@

Furthermore it is recommended to install the ontology editor 
Prot\`eg\`e\footnote{\href{http://protege.stanford.edu}{http://protege.stanford.edu}} in order to be to easily maintain the 
knowledge database of installable packages. At the time of writing, no Linux distribution is offering packages for this fine tool.

@@

The current version of the Janitor can be configured using the common file \textit{arc.conf}. It is expected in the 
configuration directory \textit{etc}. The Janitor is using the environment variable \textit{NORDUGRID\_CONFIG}
to determine the location of the corresponding file. If that variable is not set, the default location \textit{/etc/arc.conf} will be used. 
The configuration is assigned by the section \lbrack janitor\rbrack. Table~\ref{tab:arc_conf_tags} describes the available tags
for Janitor's configuration.

@@

The parameter \texttt{enable} specifies whether the Janitor shall be used within A-REX or not. Use the value \texttt{"0"} to disable Janitor. The
\texttt{uid} and the \texttt{gid} are defining which effective the user and group id shall be used for Janitor. 
The \texttt{registrationdir} describes the directory in which the subdirectories \texttt{jobs} and \texttt{rtes} will be created.
In these directories the states of the jobs and the runtime environments will be stored. 
The knowledge base of installable packages is specified by the parameter \texttt{catalog}. 
Its value can be any kind of URL pointing to a file written in the Resource Description Framework (RDF) format.
The specification of the RDF file will be explained in detail in section~\ref{sec:catalog}.
The parameter \texttt{downloaddir} assigns the directory in which the installation files will be saved after they have been 
downloaded or copied from the repository which was specified by the catalog. The \texttt{installationdir} finally specifies the
directory into which all packages will be installed and should be 
available for all computing elements, i.e.\ by using a shared volume.
If the configuration file furthermore contains the \texttt{runtimedir} tag within the section \texttt{grid-manager}, the Janitor will also
create a symbolic link in the \texttt{runtimedir} pointing to the configuration script of the installation done by the Janitor.
The tags \texttt{jobexpirytime} and \texttt{rteexpirytime} are used for automated cleanup and is defined in seconds.
The default value for the \texttt{jobexpirytime} is seven days and for the \texttt{rteexpirytime} three days. 
The additional tags \texttt{allow\_base} \texttt{deny\_base} \texttt{allow\_rte} and \texttt{deny\_rte} are used to include
or exclude certain base packages or runtime environments of the catalog. This feature is useful, if the catalog is maintained by
a higher organization.
The path to the log4perl configuration file is defined by the tag \texttt{logconf}.
An example how to configure ARC and log4perl is provided in the Listings~\ref{lst:arc_conf} and ~\ref{lst:log_conf}.

### 2009-06-08 doc/manuals/janitor/tex_introduction/introduction.tex
Michael's additions and some pending corrections by Steffen.

Janitor is a service for the automated installation of runtime environment for grid computing elements.
It is addressed transparently via the A-REX service used for job submission but can also be used as a standalone tool.

@@

A core problem remains to distribute a locally working solution, the Know-How, quickly across all
contributing sites, i. e., without manual interference. Every scientific discipline has its respective own set of
technologies for the distribution of work load. E.g. research in bioinformatics requires access to so many
different tools and databases, that few sites, if any, install them all. Instead, the use of web services became
a commodity, with all their intrinsic problems as there are bottlenecks and restrictions of repeated access.
The EU project KnowARC1\footnote{\href{http://www.knowarc.eu}{http://www.knowarc.eu}} amongst other challenges
with the here presented work extends the NorduGrid's
Advanced Research Connector (ARC) grid middleware~\cite{ELLERT_2007} towards an infrastructure for the
automated installation of software packages.

An automation of the software installation, referred to as dynamic Runtime Environments, seems the only
approach to use the computational grid to its full potential. Components of workflows shall be spawned
as jobs in a computational grid using dynamic Runtime Environments rather than as shared web services.
The grid introduces an extra level of parallelism that web services cannot provide. The required short
response times and the heterogeneous education of site-administrators on a grid demand an automatism for
the installation of software and databases without manual interference~\cite{BAYER_2007}.

@@

This document will start with a chapter on how to set-up the Janitor locally. The following chapter will then give furhter instructions on how to 
use the Janitor with A-REX and/or without A-REX. Afterwards, in the third chapter, the maintenance of the program will be presented, 
which is basically covering the method how to prepare new runtime environments. Deeper insights on the design of the Janitor will 
be given by the subsequent forth chapter. In the fifth chapter, an outlook to anticipated future developments will given.

>>>>

Janitor is a service for the automated installation of runtime environment for grid computing elements.
It is addressed transparently via the A-REX service for job submission but can also be used as a standalone tool.

@@

A core problem remains to distribute a locally working solution, the Know-How, quickly across all
contributing sites, i.e., without manual interference. Every scientific discipline has its respective own set of
technologies for the distribution of work load. E.g. research in bioinformatics requires access to so many
different tools and databases, that few sites, if any, install them all. Instead, the use of web services became
a commodity, with all their intrinsic problems as there are bottlenecks and restrictions of repeated access.
The EU project KnowARC\footnote{\href{http://www.knowarc.eu}{http://www.knowarc.eu}} amongst other challenges
with the here presented work extends the NorduGrid's
Advanced Research Connector (ARC) grid middleware~\cite{ELLERT_2007} towards an infrastructure for the
automated installation of software packages.

An automation of the software installation, referred to as dynamic Runtime Environments, seems the only
approach to use the computational grid to its full potential. Components of workflows shall be spawned
as jobs in a computational grid using dynamic Runtime Environments rather than accessing a web service
at one particular machine that is shared amongst all users.
The grid introduces an extra level of parallelism that web services cannot provide. The demands for short
response times and the heterogeneous education of site-administrators on a grid demand an automatism for
the installation of software and databases without manual interference~\cite{BAYER_2007}.

@@

This document first dedicates a chapter on how to set-up the Janitor locally. It is followed by a chapter that gives furhter instructions on how to 
use the Janitor with A-REX and/or without A-REX. Afterwards, in the third chapter, the maintenance of the program will be presented, 
which is basically covering the method how to prepare new runtime environments. Deeper insights on the design of the Janitor will 
be given by the forth chapter. The document ends with an outlook to anticipated future developments and opportunities.

\section*{Abbreviations}

\begin{tabular}{l@{--}p{150mm}}
RTE&Runtime Environment\\
dRTE&dynamic Runtime Environment\\
RDF&Resource Description Framework (supporting the RTE Catalog)\\
\end{tabular}

### 2009-06-08 doc/manuals/janitor/tex_maintenance/maintenance.tex
Michael's additions and some pending corrections by Steffen.

This chapter explains how to maintain the Catalog and the Janitor itself. 
In order to administrate the catalog, it is strongly recommended (but not ultimately required) to use the ontology editor Prot\`eg\`e. The first
section will explain how this is done.
The second section presents a detailed explanation how to create new packages for the Janitor.
In the last section, a typical use case in maintaining Janitor will be listed.

@@

The Catalog describes runtime environments and is either served through a web server or is distributed together with the Janitor as a regular file. 
It is specifed by a (Resource Description Framework) RDF file assigned to 
Janitor using the tag \texttt{catlog} within the configuration file (see~\ref{sec:janitor_configuration}). 
The format of the RDF file is defined by a RDF schema file \texttt{knowarc.rdfs}
which can be found along with an RDF example file \texttt{knowarc.rdf} in the janitor source directory \href{http://svn.nordugrid.org/repos/nordugrid/arc1/trunk/src/services/janitor/resources/catalog/}
{http://svn.nordugrid.org/repos/nordugrid/arc1/trunk/src/services/janitor/resources/catalog/}.\\

In order to adminstrate the catalog, the ontology editor Prot\`eg\`e should be used. Figure~\ref{fig:protege_example} shows the
editor while the MetaPackage \texttt{APPS/BIO/JASPAR-CORE-1.0} of the example file has been selected.
\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{images/protege_JASPAR.png}
    \mycaption{Example of a RDF catalog file as displayed in the program Prot\`eg\`e.}{}
    \label{fig:protege_example}
  \end{center}
\end{figure}
On the left side of the editor the class browser is placed. Three main classes are prepared: \texttt{MetaPackage}, \texttt{Note} and \texttt{Package}.
The \texttt{Metapackage} is a general platform independent description of a \texttt{Package}. 
It has one or more instances of the class \texttt{Package} and is
described by the subclasses of \texttt{Note}. The class \texttt{Note} has two subclasses: \texttt{BaseSystem} 
and \texttt{Tag} to describe the \texttt{MetaPackage}. The \texttt{BaseSystem} describes the Debian release a \texttt{Package} refers to
(i.e. here etch or sid), i.e. the name of a common installation or a virtual image.
The class \texttt{Tag} provides small keywords which can be assigned to \texttt{MetaPackages} such that
they can be found more easily.  \texttt{TarPackage} and \texttt{DebianPackage} or currently the only subclasses of \texttt{Package}.
They are representing the necessary information (i.e. URL or Packagename) for the installation. In order to have an overview how the 
classes are interacting with each other the Tables~\ref{tab:main_class_meta},\ref{tab:main_class_base}, \ref{tab:main_class_tag},
\ref{tab:main_class_debian} and \ref{tab:main_class_tar} are pictured.

@@

The dynamic Runtime Environments stored in the Catalog are presented on the aforementioned dedicated web
page\footnote{http://dre.knowarc.eu:8080/list.pl}. This site also links to both the formal Catalog in 
RDF syntax and an automated transformation to
HTML. The latter mimics the traditional site describing Runtime Environments in the Runtime Environment
Registry\footnote{http://gridrer.csc.fi/} in order to minimise issues with an eventual transition to the new system. 
That page collects descriptions for Runtime Environments to encourage human site administrators to install these.
This HTML page listing the manually or automatically installable REs is prepared by the script web/list.pl.
This script is meant to be run by a mod\_perl-enabled Apache. The script itself is only loosely coupled
with the Janitor. In the first lines of the script some variables specific to the site are set. To configure the
script these have to be changed~\cite[p. 9]{BAYER_2007}.

@@

At the time of writing, only the tape archive (tar) file format is accepted for dynamic Runtime Environment installation, 
a well accepted file format throughout the UNIX community. This section explains the inner structure of the tar files. 
Subdirectories are visualised in Figure~\ref{fig:tar_folder}.

@@

The tar file contains two directories, named {\tt control} and {\tt data}. Software is stored in the latter subdirectory,
while the files formally specifying how to deal with such packages are stored in the prior. Upon installation,
the content of the data directory is extracted to some directory \textdollar BAR. 
After this unpacking of the tar file, the Janitor executes the install script provided in the control directory. 
It is executed within the working directory \textdollar BAR. The job of this skript is to perform any necessary post-processing.
The Janitor stores the file \texttt{control/remove}. It will be executed in the same way as \texttt{control/install} just
before the tar-package is removed. In most cases \texttt{control/remove} will be empty.
Finally, the file \texttt{control/runtime} is sourced multiple times by the Grid Manager's job-submit script. After
installing the package, the Janitor changes all occurences of \%BASEDIR\% in the runtime script to \textdollar BAR.
To be offered to computing elements for an installation, the such prepared runtime environment must be
announced to a Catalog to which the Janitor on the computing element subscribes~\cite[p. 10]{BAYER_2007}.

@@

In order to have an impression how the tar files are created, several prototypes are provided at \task{URL required!!}.:\\

@@

\textbf{CRAN and BioConductor.org} To demonstrate the technical proximity to scientific communities that
provide packages for the Debian Linux distribution, a tool was prepared to transform Debian packages to
dynamic runtime environments\cite{MOELLER_2007}. This effort comprising more than 1700 packages and
thus also helps to analyse the scalability of the RDF-based tool for the analysis of dependencies between
projects.  \task{Why is CRAN and BioConductor.org mentioned here??}\\

@@

\subsection{Debian Etch}

These kind of packages are not yet supported.

\subsection{Debian Sid}

These kind of packages are not yet supported.

\subsection{Adminstrating the Catalog}

>>>>

This chapter explains how to maintain the {\em Catalog} and the {\em Janitor} itself. 
It gives detailed instructions how to create new packages for the Janitor.
To administrate the catalog, it is strongly recommended (but not ultimately required) to use the ontology editor Prot\`eg\`e,
as explained in the next section.
In the last section, a typical use case in maintaining the Janitor will be shown.

@@

The Catalog describes runtime environments and is either served through a web server or is distributed together with the Janitor as a regular file. 
It is specifed by a (Resource Description Framework) RDF file assigned to 
Janitor using the tag \texttt{catalog} within the configuration file (see~\ref{sec:janitor_configuration}). 
The format of the RDF file is defined by a RDF schema file \texttt{knowarc.rdfs}
which can be found along with an RDF example file \texttt{knowarc.rdf} in the janitor source directory \href{http://svn.nordugrid.org/repos/nordugrid/arc1/trunk/src/services/janitor/resources/catalog/}
{http://svn.nordugrid.org/repos/nordugrid/arc1/trunk/src/services/janitor/resources/catalog/}.\\

In order to adminstrate the catalog, the ontology editor Prot\`eg\`e should be used. Figure~\ref{fig:protege_example} shows the
editor while the MetaPackage \texttt{APPS/BIO/JASPAR-CORE-1.0} of the example file has been selected.
\begin{figure}
  \begin{center}
    \includegraphics[width=\textwidth]{images/protege_JASPAR.png}
    \mycaption{Example of a RDF catalog file as displayed in the program Prot\`eg\`e.}{}
    \label{fig:protege_example}
  \end{center}
\end{figure}
On the left side of the editor the class browser is placed. Three main classes are prepared: \texttt{MetaPackage}, \texttt{Note} and \texttt{Package}.
The \texttt{Metapackage} is a general platform independent description of a \texttt{Package}. 
It has one or more instances of the class \texttt{Package} and is
described by the subclasses of \texttt{Note}. The class \texttt{Note} has two subclasses: \texttt{BaseSystem} 
and \texttt{Tag} to describe the \texttt{MetaPackage}. The \texttt{BaseSystem} describes the Debian release a \texttt{Package} refers to
(i.e. here etch or sid), i.e. the name of a common installation or a virtual image.
The class \texttt{Tag} provides small keywords which can be assigned to \texttt{MetaPackages} such that
they can be found more easily.  \texttt{TarPackage} and \texttt{DebianPackage} or currently the only subclasses of \texttt{Package}.
They are representing the necessary information (i.e. URL or Packagename) for the installation. In order to have an overview how the 
classes are interacting with each other the tables~\ref{tab:main_class_meta}, \ref{tab:main_class_base}, \ref{tab:main_class_tag},
\ref{tab:main_class_debian} and \ref{tab:main_class_tar} are pictured.

@@

The dynamic Runtime Environments stored in the Catalog are presented on the aforementioned dedicated web
page\footnote{http://dre.knowarc.eu:8080/list.pl}. This site also links to both the formal Catalog in 
RDF syntax and its automated transformation to
HTML. The latter mimics the traditional site describing Runtime Environments in the Runtime Environment
Registry\footnote{http://gridrer.csc.fi/} in order to minimise issues with an eventual transition to the new system. 
That page collects descriptions for Runtime Environments to encourage human site administrators to install these.
This HTML page listing the manually or automatically installable REs is prepared by the script web/list.pl.
This script is meant to be run by a mod\_perl-enabled Apache. The script itself does not contribute to the 
core functionality of the Janitor.
In the first lines of the script some variables specific to the site are set. To configure the
script these have to be changed~\cite[p. 9]{BAYER_2007}.

@@

At the time of writing, only the tape archive (tar) file format is accepted for dynamic Runtime Environment installation, 
a well accepted file format throughout the UNIX community. The reflects the traditional manual approach towards RTE in ARC,
for which one directory is made available to all compute nodes. This section explains the inner structure of the tar files. 
Subdirectories are visualised in Figure~\ref{fig:tar_folder}.

@@

The tar file contains two directories, named {\tt control} and {\tt data}. Software is stored in the latter subdirectory,
while the files formally specifying how to deal with such packages are stored in the prior. Upon installation,
the content of the data directory is extracted to some directory \textdollar BAR. 
After this unpacking of the tar file, the Janitor executes the install script provided in the control directory. 
It is executed within the working directory \textdollar BAR. The job of this skript is to perform any necessary post-processing.
The Janitor stores the file \texttt{control/remove}. It will be executed in the same way as \texttt{control/install} just
before the tar-package is removed. In most cases \texttt{control/remove} will be empty.
Finally, the file \texttt{control/runtime} is sourced multiple times by the Grid Manager's job-submit script. After
installing the package, the Janitor changes all occurences of \%BASEDIR\% in the runtime script to \textdollar BAR.
Once the tar file was prepared, it must its entry to a RTE Catalogue~\cite[p. 10]{BAYER_2007}.

From such Catalogues, the Janitor finds all information to install packages that possibly have never been installed
on the site before. The offers of RTEs in a catalogue are cross-checked against the local infrastructure and a subset
of the available packages will be accepted as "installable". This list of installable RTEs is forwarded to the
grid information system.

@@

In order to have an impression how the tar files are created, several
prototypes are provided at \url{http://dre.knowarc.eu}.

@@

%\textbf{CRAN and BioConductor.org} To demonstrate the technical proximity to scientific communities that
%provide packages for the Debian Linux distribution, a tool was prepared to transform Debian packages to
%dynamic runtime environments\cite{MOELLER_2007}. This effort comprising more than 1700 packages and
%thus also helps to analyse the scalability of the RDF-based tool for the analysis of dependencies between
%projects.  \task{Why is CRAN and BioConductor.org mentioned here??}\\

@@

%\subsection{Debian Etch}
%
%These kind of packages are not yet supported.
%
%\subsection{Debian Sid}
%
%These kind of packages are not yet supported.
%
%\subsection{Adminstrating the Catalog}

### 2009-06-08 doc/manuals/janitor/tex_usage/usage.tex
Michael's additions and some pending corrections by Steffen.

\lstsetXRSL
\begin{lstlisting}[
        label=lst:xrsl_job,
        caption={ [Job submission using the xRSL job description language.]
                  \textbf{Job submission using the xRSL job description language.}}
        ]
 &
 (executable = "run.sh" )
 (arguments = "weka.classifiers.trees.J48" "-t" "weather.arff")
 ("inputfiles" = ("weather.arff" "" ))
 ("stderr" = "stderr" )
 ("stdout" = "stdout" )
 ("gmlog" = "gmlog" )
 ("runtimeenvironment" = "APPS/BIO/WEKA-3.4.10")
 ("runtimeenvironment" = "APPS/BIO/WISE-2.4.1-5")
\end{lstlisting}

@@

In addition to using the Janitor together with A-REX, the Janitor can also be used as a standalone commandline tool.
The available commands are listed in the Table~\ref{janitor_commandline_man}.
\begin{table}[!h]
   \begin{center}
        \mycaption{Overview about the available commands in Janitor.}{}
        \label{tab:janitor_commandline_man}
	\begin{tabular}{p{0.5cm}p{2cm}p{11cm}}
	\multicolumn{3}{l}{\textbf{janitor [COMMAND] [JOB-ID] [RTE] \dots}} \\
	\multicolumn{3}{l}{\textbf{Command:}}\\
	&	register			& Registers a job and a set of runtime environments in the Janitor database. Requires the parameters [JOB-ID] and a list of [RTE]s.\\
	&	deploy				& Downloads and installs the desired runtime environments. Requires the name of an already registered [JOB-ID].\\
	&	remove				& Removes the placeholder of the job on the runtime environments. If no more jobs are using the runtime environment and the lifespan of the runtime environment has be expired, the runtime environment can be removed using the \texttt{sweep} command. Requires the [JOB-ID] to be removed.\\
	&					&\\
	&	sweep				& Removes unused runtime environments. No further arguements are required. Using the option \texttt{--force} enforces the removal of all unused runtime environments. Runtime environments having the state FAILED will not be removed.\\
	&	setstate			& Changes the state of a dynamically installed runtime environment. This might be useful in case a runtime environment with a state FAILED shall be removed (new state might be REMOVAL\_PENDING). Requires the argument [STATE] followd by a list of [RTE]s.\\
	&					&\\
	&	search				& Performs a simple search in the catalog and the manually installed runtime environments (\texttt{runtimedir}). Requires no [JOB-ID] nor [RTE]s, but only a list of string to be searched for.\\
	&	list				& Lists all information about jobs, automatically installed runtime environments and manually installed runtime environments. No additional parameters have to be passed.\\
	&	info				& Renders information about a job. Requires the parameter [JOB-ID].\\
	\multicolumn{3}{l}{\textbf{Job id:}}\\
	&					& A unique sequence of numbers. Once Janitor registered a job id it cannot register a second job having the same job id.\\
	\multicolumn{3}{l}{\textbf{Runtime environments:}}\\
	&					& Runtime environments are defined by a continuous string. The name of valid runtime environment names can be investigated using the \texttt{list} or the \texttt{search} commands. They are defined in the catalog or by the directories and scripts of the \texttt{runtimedir} of the \texttt{grid-manager}.\\
	\end{tabular} 
   \end{center}
\end{table}
The most important commands for the Janitor are \texttt{register}, \texttt{deploy} and \texttt{remove}. To 
register a job along with a set of runtime environments in Janitor, the first command \texttt{register} followed
by a job identifier and a list of runtime environments has to be used.
A job is identified by a sequence of numbers. Runtime environments are specified by a string containing the name as it is defined 
within the Catalog (resp. the runtime directory of the grid-manager).
The command \texttt{deploy} extracts the necessary dependencies of the desired runtime environments to then download and install
the required packages. %How dependecies can are defined will be explained later in the section~\ref{sec:catalog}.

% PROCEED HERE!!!!!!!! !! ! ! !
In order to remove jobs registered in Janitor the command \texttt{remove} has to be used. 
The command only removes the job entry and the lock on the runtime environment. If there are no more locks on the runtime environment
it might be deleted for real.

Easy commandline examples are provided in Listing~\ref{lst:janitor_example}.

>>>>

\lstsetXRSL
\begin{lstlisting}[
        label=lst:xrsl_job,
        caption={ [Job submission using the xRSL job description language.]
                  \textbf{Job submission using the xRSL job description language.}}
        ]
 &
 (executable = "run.sh" )
 (arguments = "weka.classifiers.trees.J48" "-t" "weather.arff")
 ("inputfiles" = ("weather.arff" "" ))
 ("stderr" = "stderr" )
 ("stdout" = "stdout" )
 ("gmlog" = "gmlog" )
 ("runtimeenvironment" = "APPS/BIO/WEKA-3.4.10")
 ("runtimeenvironment" = "APPS/BIO/WISE-2.4.1-5")
\end{lstlisting}
\task{The runtime environment names are composed out a directory name, the package name and the version number.}

@@

In addition to using the Janitor together with A-REX, the Janitor can also be used as a standalone command line tool.
The available commands are listed in table~\ref{janitor_commandline_man}.
\begin{table}[!h]
   \begin{center}
        \mycaption{Overview about the available commands in Janitor.}{}
        \label{tab:janitor_commandline_man}
	\begin{tabular}{p{0.5cm}p{2cm}p{11cm}}
	\multicolumn{3}{l}{\textbf{janitor [COMMAND] [JOB-ID] [RTE] \dots}} \\
	\multicolumn{3}{l}{\textbf{Command:}}\\
	&	register			& Registers a job and a set of runtime environments in the Janitor database. Requires the parameters [JOB-ID] and a list of [RTE]s.\\
	&	deploy				& Downloads and installs the desired runtime environments. Requires the name of an already registered [JOB-ID].\\
	&	remove				& Removes the placeholder of the job on the runtime environments. If no more jobs are using the runtime environment and the lifespan of the runtime environment has be expired, the runtime environment can be removed using the \texttt{sweep} command. Requires the [JOB-ID] to be removed.\\
	&					&\\
	&	sweep				& Removes unused runtime environments. No further arguements are required. Using the option \texttt{--force} enforces the removal of all unused runtime environments. Runtime environments having the state FAILED will not be removed.\\
	&	setstate			& Changes the state of a dynamically installed runtime environment. This might be useful in case a runtime environment with a state FAILED shall be removed (new state might be REMOVAL\_PENDING). Requires the argument [STATE] followd by a list of [RTE]s.\\
	&					&\\
	&	search				& Performs a simple search in the catalog and the manually installed runtime environments (\texttt{runtimedir}). Requires no [JOB-ID] nor [RTE]s, but only a list of string to be searched for.\\
	&	list				& Lists all information about jobs, automatically installed runtime environments and manually installed runtime environments. No additional parameters have to be passed.\\
	&	info				& Renders information about a job. Requires the parameter [JOB-ID].\\
	\multicolumn{3}{l}{\textbf{Job id:}}\\
	&					& A unique sequence of numbers. Once Janitor registered a job id, it cannot register a second job having the same job id.\\
	\multicolumn{3}{l}{\textbf{Runtime environments:}}\\
	&					& Runtime environments are defined by a continuous string. The name of valid runtime environment names can be investigated using the \texttt{list} or the \texttt{search} commands. They are defined in the catalog or by the directories and scripts of the \texttt{runtimedir} of the \texttt{grid-manager}.\\
	\end{tabular} 
   \end{center}
\end{table}
The most important commands for the Janitor are \texttt{register}, \texttt{deploy} and \texttt{remove}. To 
register a job along with a set of runtime environments in Janitor, the first command \texttt{register} followed
by a job identifier and a list of runtime environments has to be used.
A job is identified by a sequence of numbers. Runtime environments are specified by a string containing the name as it is defined 
within the Catalog (resp. the runtime directory of the grid-manager).
The command \texttt{deploy} extracts the necessary dependencies of the desired runtime environments to then download and install
the required packages. %How dependecies can are defined will be explained later in the section~\ref{sec:catalog}.

% PROCEED HERE!!!!!!!! !! ! ! !
In order to remove jobs registered in Janitor, the command \texttt{remove}
has to be used.  The command only removes the job entry and the lock
on the runtime environment. If there are no more locks on the runtime
environment it might be deleted for real. The demand to pass a job
number for the removal of a RTE is irritating at first.  This shall
prevent the removal of runtime envrironments that are still being used
by jobs in the system. Instead, the janitor is informed about a job's
termination and is requested to remove the assignment of that job to the
runtime environment. Only those RTEs with no job-assignment are eligible
for being sweeped. RTEs come with an expiry time or the command may be
performed via the command line.

Easy command line examples are provided in Listing~\ref{lst:janitor_example}.
